---
title: "Marco_rmd"
author: "Marco Lee"
date: "March 26, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Load library and wrangled data
```{r}
library(dplyr)
library(tidyverse)
library(seqinr)
library(BiocManager)
library(Biostrings)
library(rBLAST)
load("MeltedData.RData") #load wrangled data
```

#Load Manduca sexta protein sequences
```{r}
msex.prot<-readAAStringSet("ms_ogs_proteins.fasta")
names(msex.prot)<-gsub("(Msex2.)(\\d+).+", "\\1\\2", names(msex.prot)) #remove additional comments after gene ID
```

#Identify the three most differentially expressed genes in each of the three tissues
```{r}
headdat2$gene_id<- as.character(headdat2$gene_id)
fatdat2$gene_id<- as.character(fatdat2$gene_id)
gutdat2$gene_id<- as.character(gutdat2$gene_id)

goi<- c(headdat2$gene_id[1:3], gutdat2$gene_id[1:3], fatdat2$gene_id[1:3])
goi<-unique(goi)
goi<-data.frame(goi=goi) #six unique genes
```

#Select the protein sequences of the 6 genes of interest from msex.prot
```{r}
msex.prot.goi<-msex.prot[goi$goi]; rm(msex.prot)
msex.prot.goi
```

#BLAST
##this is my local directory specifically for NCBI BLAST
```{r}
setwd("D:/user/Documents/NCBI_BLAST")
```

##Make blast database
Databases are too big to upload onto GitHub. They are available upon request.
```{r}
makeblastdb("insecta_nt.fasta", dbtype="nucl", args="-out insecta_nt") #create blast database of insecta nucleotides
makeblastdb("insecta_prot.fasta", dbtype="prot", args="-out insecta_prot") #create blast database of insecta proteins
```

##tblastn with Insecta nucleotide
```{r}
db.nt<-blast(db="insecta_nt", type="tblastn")
db.nt

goi1.n<-predict(db.nt, msex.prot.goi[1])
goi1.n[1:5,]

goi2.n<-predict(db.nt, msex.prot.goi[2])
goi2.n[1:5,]

goi3.n<-predict(db.nt, msex.prot.goi[3])
goi3.n[1:5,]

goi4.n<-predict(db.nt, msex.prot.goi[4])
goi4.n[1:5,]

goi5.n<-predict(db.nt, msex.prot.goi[5])
goi5.n[1:5,]

goi6.n<-predict(db.nt, msex.prot.goi[6])
goi6.n[1:5,]

goi.n<-rbind(goi1.n[1:5,], goi2.n[1:5,], goi3.n[1:5,], goi4.n[1:5,], goi5.n[1:5,], goi6.n[1:5,])
goi.n$SubjectID<-as.character(goi.n$SubjectID)
goi.n<-goi.n[complete.cases(goi.n),]
str(goi.n)
```

##blastp with Insecta proteins
```{r}
db.p<-blast(db="insecta_prot", type="blastp")
db.p

goi1.p<-predict(db.p, msex.prot.goi[1])
goi1.p[1:5,]

goi2.p<-predict(db.p, msex.prot.goi[2])
goi2.p[1:5,]

goi3.p<-predict(db.p, msex.prot.goi[3])
goi3.p[1:5,]

goi4.p<-predict(db.p, msex.prot.goi[4])
goi4.p[1:5,]

goi5.p<-predict(db.p, msex.prot.goi[5])
goi5.p[1:5,]

goi6.p<-predict(db.p, msex.prot.goi[6])
goi6.p[1:5,]

goi.p<-rbind(goi1.p[1:5,], goi2.p[1:5,], goi3.p[1:5,], goi4.p[1:5,], goi5.p[1:5,], goi6.p[1:5,])
goi.p$SubjectID<-as.character(goi.p$SubjectID)
goi.p<-goi.p[complete.cases(goi.p),]
```
I look through the top 5 hits of each BLAST (total of 10 over tblastn and blastp) for each gene to identify the consensus and most likely identity of each gene.

#Export result
```{r}
save.image("msex_goi.RData")
```
